---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r q1, echo=TRUE}
data<-read.csv("activity.csv", stringsAsFactors = FALSE)
data$date<-as.Date(data$date)
```


## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r q2_1, echo=TRUE, message=FALSE}
  require(dplyr)
  
  by_day<-group_by(na.omit(data), date)
  steps_by_day<-summarise(by_day,totalSteps=sum(steps))
  summary(steps_by_day)
```
2. Make a histogram of the total number of steps taken each day
```{r q2_2, echo=TRUE, message=FALSE}
  require(ggplot2)
  
  p<-ggplot(data = steps_by_day,aes(x = totalSteps)) +
      geom_histogram(binwidth = 5000)+
      labs(title = " Total Number of Steps Taken Per Day (NA removed)")+
      theme(plot.title = element_text(colour = "blue"))
  print(p)
```
3. Calculate and report the mean and median of the total number of steps taken per day
```{r q2_3, echo=TRUE, message=FALSE}
  mean<-mean(steps_by_day$totalSteps)
  median<-median(steps_by_day$totalSteps)
```
####-mean of the total number of steps taken per day:`r mean`
####-median of the total number of steps taken per day:`r median`

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r q3_1, echo=TRUE, message=FALSE}
  by_interval<-group_by(data, interval)
  steps_by_interval<-summarise(by_interval,avgSteps=mean(steps,na.rm = TRUE))
  p<-ggplot(data = steps_by_interval,aes(x = interval, y = avgSteps)) +
    geom_line()
  print(p)
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r q3_2, echo=TRUE, message=FALSE}
  maxAvgSteps<-max(steps_by_interval$avgSteps)
  p<-p+
    geom_point(data = filter(steps_by_interval, avgSteps==maxAvgSteps),colour="red")+
    geom_text(aes(label=interval),data = filter(steps_by_interval, avgSteps==maxAvgSteps),nudge_y = 10)
  print(p)
  steps_by_interval<-arrange(steps_by_interval, desc(avgSteps))
  i<-1
  for(i in 1:nrow(steps_by_interval))
  {
    if(steps_by_interval$avgSteps[i+1]<steps_by_interval$avgSteps[i]){
      break
    }
  }
  maxInterval<-as.character(steps_by_interval$interval[1])
  if(i>1){
    for(j in 2:i){
      maxInterval<-paste(maxInterval, ",", steps_by_interval$interval[i])
    }
  }
```
Interval `r maxInterval` contains the maximum number of steps.

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset 
```{r q4_1, echo=TRUE, message=FALSE}
  naRows<-nrow(data[is.na(data$steps),])
```
Dataset contain `r naRows` missing values.

2. Devise a strategy for filling in all of the missing values in the dataset 
**Use mean of each interval to fill NA**

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r q4_3, echo=TRUE, message=FALSE}
  ##Use mean of each interval to fill NA
  by_interval<-group_by(data,interval)
    steps_by_interval<-summarise(by_interval,avgSteps=mean(steps,na.rm = TRUE))
    na_interval<-distinct(data[is.na(data$steps),],interval)
    for (i in 1:nrow(na_interval)){
      data$steps[is.na(data$steps)&data$interval==na_interval$interval]<-steps_by_interval$avgSteps[steps_by_interval$interval==na_interval$interval]
    }
  naRows<-nrow(data[is.na(data$steps),])
  str(data)
```
Dataset contain `r naRows` missing values.

4. Make a histogram of the total number of steps taken each day
```{r q4_4, echo=TRUE, message=FALSE}
  by_day<-group_by(data, date)
  steps_by_day<-summarise(by_day,totalSteps=sum(steps,na.rm = TRUE))
  p<-ggplot(data = steps_by_day,aes(x = totalSteps)) +
      geom_histogram(binwidth = 5000)+
      labs(title = " Total Number of Steps Taken Per Day (NA filled)")+
      theme(plot.title = element_text(colour = "blue"))
  print(p)
  mean2<-mean(steps_by_day$totalSteps)
  median2<-median(steps_by_day$totalSteps)
```
####-mean of the total number of steps taken per day:`r mean`, differ: `r mean2-mean`
####-median of the total number of steps taken per day:`r median`, differ: `r median2-median`

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels ¨C ¡°weekday¡± and ¡°weekend¡±
```{r q5_1, echo=TRUE, message=FALSE}
  getWeek<-function(date=as.Date()){
    w<-as.character()
    for (i in 1:length(date)){
      if(weekdays(date[i])=="ÐÇÆÚÁù"|weekdays(date[i])=="ÐÇÆÚÈÕ"){
        w<-c(w,"weekend")
      }
      else{
        w<-c(w,"weekday")
      }
    }
    w
  }
  data<-mutate(data, week=as.factor(getWeek(date)))
  head(data,5)
```
2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r q5_2, echo=TRUE, message=FALSE}
  by_interval<-group_by(data, week, interval)
  steps_by_interval<-summarise(by_interval,avgSteps=mean(steps,na.rm = TRUE))
  
  p<-ggplot(data = steps_by_interval, aes(x = interval, y = avgSteps))+
    geom_line()+
    facet_grid(facets = week~.)
  print(p)
```
